---
title: "Species_interaction_analysis"
output: html_document
date: "2025-01-27"
---

```{r setup, include=FALSE}
rm(list=ls())
library("ggplot2")
library("tidyverse")
library("gridExtra")
#library(rEDM)
library("parallel")
library("pracma")
library("purrr")
library("signal")
library("ggfortify")
library("data.table")
library("patchwork")
library("codyn")
library("ggrepel")
library("lme4")
library("lmerTest")
library("MuMIn")
library("RColorBrewer")
library("broom")
library("relaimpo")
library("lavaanPlot")
library(ggbeeswarm)
library(performance)
library(broom)
library(kableExtra)
library(gt)
library(lavaan)
#devtools::install_github("canankarakoc/r_package_EDMhelper/EDMhelper")
```

# Load data sets
```{r cars}
load("Data/complete_ts.RData")
load("Data/pop_aggr.RData")
interactions_df <- read_csv("Data/interactions_df_test.csv")


```
# Get time of extinction 

Coefficients were only considered for time points where species biomass did not drop below 1/10 of initial biomass
```{r pressure, echo=FALSE}
#### Remove coefficients of temperature

interactions_df$day[interactions_df$day==0.00]<-60.28

interactions_df<-dplyr::filter(interactions_df,coefficient_name!="∂C/∂norm_sampleT" & coefficient_name!="∂D/∂norm_sampleT" & coefficient_name!="∂L/∂norm_sampleT" &
                          coefficient_name!="∂S/∂norm_sampleT" & coefficient_name!="∂P/∂norm_sampleT")

#### Get extinctions of species


biom_day_1<-complete_ts%>%dplyr::filter(day==1)%>%group_by(sample_ID,predict_spec)%>%mutate(biomass_day1=biomass)

ext_ts<-complete_ts%>%inner_join(biom_day_1%>%dplyr::select(sample_ID,predict_spec,biomass_day1),by=c("sample_ID","predict_spec"))



ext_ts <- ext_ts %>%
  group_by( sample_ID,predict_spec) %>%
  arrange(day) %>%
  mutate(
    # Create a flag for non-zero tot_biomass; species below 1/10 as threshold
    non_zero_biomass = biomass >=  biomass_day1 / 10,
    # Find the last non-zero day
    last_non_zero_day = ifelse(any(non_zero_biomass), max(day[non_zero_biomass], na.rm = TRUE), NA),
    # Find the first zero day after the last non-zero day
    ext = ifelse(
      !is.na(last_non_zero_day),
      min(day[!non_zero_biomass & day > last_non_zero_day], na.rm = TRUE),
      NA
    ),
    # Flag for days after the first zero day
    is_after_first_zero_day = ifelse(!is.na(ext), day > ext, FALSE)
  ) %>%
  ungroup()

ext_ts<-ext_ts%>%mutate(rep_var=sub("_[^_]+$", "", sample_ID))



df_first_zero<-ext_ts%>%group_by(sample_ID,predict_spec)%>%summarize(ext=unique(ext))

#### Create interactor variable

interactions_df<-interactions_df%>%group_by(coefficient_name)%>%mutate(predict_spec=strsplit(coefficient_name,split="")[[1]][nchar(coefficient_name)])

interactions_df<-interactions_df%>%inner_join(df_first_zero,by=c("sample_ID","predict_spec"))


#### Aggregate interactions,  only using time series before extinction for each interactor

int_aggr<-interactions_df%>%group_by(theta,sample_ID)%>%dplyr::filter(day<=ext)%>%
  summarize(
            mean_interaction=mean(coefficient_value,na.rm=T))
int_aggr$theta[is.na(int_aggr$theta)]="none"
int_aggr<-int_aggr%>%dplyr::filter(theta%in%c("var","none"))

int_aggr<-int_aggr%>%group_by(theta)%>%full_join(pop_aggr,join_by(sample_ID))%>%ungroup()


int_aggr$mean_interaction[is.na(int_aggr$mean_interaction)]=0


```


# Get average interaction coefficient
```{r}


bootstrap_mean <- function(data, n_boot = 1000) {
  boot_means <- numeric(n_boot)
  for (i in seq_len(n_boot)) {
    resample <- sample(data, size = length(data), replace = TRUE)
    boot_means[i] <- mean(resample)
  }
  return(boot_means)
}

set.seed(42)
boot_test<-bootstrap_mean(int_aggr$mean_interaction,n_boot=1000)

ci <- quantile(boot_test, probs = c(0.025, 0.975))
cat("Bootstrap 95% CI:", ci, "\n")

mean(int_aggr$mean_interaction)

```


# Get residual temperature and nutrients
```{r}
int_aggr2<- int_aggr %>%
  # Remove the units from the 'nutrients' and 'temperature' columns
  mutate(
    nutrients = as.numeric(gsub(" g/L", "", nutrients)),  # Convert nutrients to numeric
    temperature = gsub(" °C", "", temperature)            # Remove the unit but keep as character
  ) %>%
  # Convert temperature ranges to numeric codes using case_when
  mutate(
    temperature = case_when(
      temperature == "18-21" ~ 1,
      temperature == "22-25" ~ 2,
      temperature == "25-28" ~ 3,
      TRUE ~ NA_real_         # Handle unexpected values with NA
    )
  )

lm_temp_resid<-lm(data=int_aggr2, temperature ~ log10(balance_f))

lm_nut_resid<-lm(data=int_aggr2, nutrients ~ log10(balance_f))
 
resid_temp<-lm_temp_resid$residuals

resid_nut<-lm_nut_resid$residuals
 
int_aggr2<-int_aggr2%>%mutate(resid_temp=resid_temp,resid_nut=resid_nut)
 
```

# Linear mixed effect model

The same model as in the main text, but with interaction coefficient as an additional explanatory variable. Negative interaction coefficients indicate competition while positive indicate fascilitation.
```{r}
model_int<- lmer(log10(stability) ~ log10(balance_f) +  richness + resid_temp * resid_nut+ mean_interaction+
                        (1 | composition), data = int_aggr2, REML = FALSE)
```


```{r}


tbl_summary <- gtsummary::tbl_regression(model_int, intercept = TRUE,
                                          label = list("(Intercept)" = "Intercept",
                                                       log_balance = "log10(balance)",
                                                       nut_scaled = "nutrients (scaled)",
                                                       temp_scaled = "temperature (scaled)",
                                                       "nut_scaled*temp_scaled" = "nutrients x temperature (scaled)"),
                                          add_estimate_to_reference_row = TRUE,  # Adds reference values for categorical predictors
                                          add_pairwise_contrasts = TRUE,
                                          add_significance_stars = TRUE) %>%
  gtsummary::modify_table_body(
    ~ .x %>%
      dplyr::mutate(
        t_value = ifelse(!is.na(estimate), round(statistic, 2), NA),  # Use 'statistic' for t-value
        # Remove std_error, as it is no longer needed
        std_error = NULL
      )
  ) %>%
  # Reorder columns: estimate, 95% CI, t-value, p-value
  gtsummary::modify_header(label = "**Predictor**", 
                           estimate = "**Estimate**", 
                           conf.int = "**95% CI**", 
                           t_value = "**t-value**", 
                           p.value = "**p-value**") %>%
  gtsummary::modify_footnote(everything() ~ NA) %>%  # Remove default footnotes for clean look
  gtsummary::bold_labels() %>%
  gtsummary::as_gt() %>%
  tab_header(title = "Linear Regression Results") %>%
  # Ensure p-value is the last column
  gt::cols_move(p.value, after = last_col()) %>%
  # Bold significant p-values
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = vars(p.value),
      rows = p.value < 0.05
    )
  )


tbl_summary
```


Mean interaction coefficient is not significant (est. = 0.03, p=0.2)


```{r}
sem_aggr <- int_aggr2 %>%
  ungroup() %>%  # Ensure there is no grouping
  mutate(
    log_balance_f = log10(balance_f),
    stability = log10(1 / CV),
    richness = as.numeric(richness),
    temperature=temperature,
    asynchrony_Gross= (-synchrony_Gross),
    pop_stability= log10(1/w_pop_CV),
    pop_var = log10(pop_V)
    #Keep it as an ordered factor
  )

model1 <- '
  stability ~ asynchrony_Gross
  +pop_stability 
 
  asynchrony_Gross ~ log_balance_f + nutrients +mean_interaction
  pop_stability~log_balance_f + nutrients +mean_interaction
  
  

'




# Fit the model

fit1 <- sem(model1,estimator="MLM",meanstructure = TRUE,data = sem_aggr%>%dplyr::filter(!is.na(asynchrony_Gross)))
#modificationindices(fit1C)

# Summarize the results
summary(fit1, standardized = TRUE,rsquare=T, fit.measures = TRUE)
```




