---
title: "When will asynchrony be more important than population stability?"
author: "Til Hämmig, Francesco Polazzo, Owen L. Petchey, Frank Pennekamp"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    code_folding: hide
    keep_md: yes
    fig_caption: true  
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
rm(list=ls())
library("ggplot2")
library("tidyverse")
library("gridExtra")
#library(rEDM)
library("parallel")
library("pracma")
library("purrr")
library("signal")
library("ggfortify")
library("data.table")
library("patchwork")
library("codyn")
library("ggrepel")
library("lme4")
library("lmerTest")
library("MuMIn")
library("RColorBrewer")
library("broom")
library("relaimpo")
library("lavaanPlot")
library(ggbeeswarm)
library(performance)
library(broom)
library(kableExtra)
library(gt)
library(lavaan)
#devtools::install_github("canankarakoc/r_package_EDMhelper/EDMhelper")

```

```{r include = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      include = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE)
```



# When will asynchrony be more important than population stability?
```{r}
source("r/our_functions.r")
## Create the communities
num_spp <- c(2, 4, 8, 16)
max_minus_min <- round(seq(1, 50, length = 45),3)
mean <- round(seq(-25, 25, length = 49),3)
rd_vals <- function(num_spp, max_minus_min, mean) {
  seq(mean-max_minus_min/2, mean+max_minus_min/2, length = num_spp)
}
dd <- crossing(num_spp,
               max_minus_min,
               mean) %>%
  rowwise() %>%
  mutate(rd_vals = list(rd_vals(num_spp, max_minus_min, mean))) %>%
  unnest(cols = rd_vals)
```

```{r}
## calculate the diversity metrics
summary_stats <- dd %>%
  group_by(num_spp, max_minus_min, mean) %>%
  summarise(dissimilarity = resp_div(rd_vals,sign_sens = FALSE),
            divergence = resp_div(rd_vals,sign_sens = TRUE),
            sign = sign(rd_vals),
            balance = abs(sum(rd_vals)),
            magnitude = sum(abs(rd_vals)),
            symmetry = mean(sign)) %>%
  pivot_longer(names_to = "variable", values_to = "value", 4:5)


```


```{r}
# plot summary stats balance vs magnitude by num_spp
summary_stats %>%
  ggplot(aes(x = balance, y = magnitude)) +
  geom_point() +
  facet_wrap( ~ num_spp, scales = "free") +
  xlab("Balance") +
  ylab("Magnitude")

# create a new column called range with the range of magnitude values for each balance value
summary_stats <- summary_stats %>%
  group_by(num_spp, balance) %>%
  mutate(range = max(magnitude) - min(magnitude),
         mean = mean(magnitude))

library(scales)
# modify the above plot so that the gradient is present in each panel
summary_stats %>%
  group_by(num_spp) %>%
  mutate(magnitude_rescaled = rescale(balance / magnitude)) %>% # Rescale within each group
  ggplot(aes(x = balance, y = magnitude, color = magnitude_rescaled)) +
  geom_point() +
  facet_wrap(~num_spp, scales = "free") +
  xlab("Balance") +
  ylab("Magnitude") +
  scale_color_viridis_c(option = "plasma") +
  labs(color = "Community Imbalance", x = "Balance") +
  theme_bw()

```


```{r}
library(scales)
# modify the above plot so that the gradient is present in each panel
summary_stats %>%
  group_by(num_spp) %>%
  mutate(magnitude_rescaled = rescale(balance / num_spp)) %>% # Rescale within each group
  ggplot(aes(x = symmetry, y = magnitude, color = magnitude_rescaled)) +
  geom_point() +
  facet_wrap(~num_spp, scales = "free") +
  xlab("Balance") +
  ylab("Magnitude") +
  scale_color_viridis_c(option = "plasma") +
  labs(color = "Balance", x = "Symmetry") +
  theme_bw()

```


```{r}
library(ggplot2)
library(dplyr)
library(scales)

# Create the original plot
summary_stats <- dd %>%
  group_by(num_spp, max_minus_min, mean) %>%
  summarise(
    dissimilarity = resp_div(rd_vals, sign_sens = FALSE),
    divergence = resp_div(rd_vals, sign_sens = TRUE),
    sign = sign(rd_vals),
    balance = abs(sum(rd_vals)),
    magnitude = sum(abs(rd_vals)),
    symmetry = mean(sign)
  ) %>%
  pivot_longer(names_to = "variable", values_to = "value", 4:5)

# Plot: Relationship between magnitude, balance, and symmetry
summary_stats %>%
  group_by(num_spp) %>%
  mutate(magnitude_rescaled = rescale(balance / num_spp)) %>%
  ggplot(aes(x = symmetry, y = magnitude, color = magnitude_rescaled)) +
  geom_jitter(width = 0.05) +
  facet_wrap(~num_spp, scales = "free") +
  scale_color_viridis_c(option = "plasma") +
  labs(color = "Rescaled Balance", x = "Symmetry", y = "Magnitude") +
  theme_bw() +
  ggtitle("Original Data: Magnitude vs Symmetry across num_spp")

```

Now, let's generate rd_vals using different distributions to see if the patterns hold. We'll consider:

Normal Distribution (rnorm).
Uniform Distribution (runif).
Skewed Distribution (e.g., using a gamma distribution rgamma).


**Comparing Across Distributions**:
For normal distributions, the pattern of high magnitude corresponding with high balance at low num_spp values (like num_spp = 2) is expected due to the symmetry around the mean.
For uniform distributions, the spread of rd_vals is even, so the same relationship is expected. However, the range of values is fixed, which may affect how often high magnitude coincides with high balance at higher num_spp.
For gamma (skewed) distributions, we may see different patterns. Gamma distributions are positively skewed, leading to asymmetry in rd_vals, potentially increasing the variance in magnitude and balance at all levels of num_spp.

```{r}
# Function to generate rd_vals from different distributions
generate_rd_vals <- function(num_spp, max_minus_min, mean, dist = "normal") {
  if (dist == "normal") {
    rd_vals <- rnorm(num_spp, mean = mean, sd = max_minus_min / 4)
  } else if (dist == "uniform") {
    rd_vals <- runif(num_spp, min = mean - max_minus_min / 2, max = mean + max_minus_min / 2)
  } else if (dist == "gamma") {
    rd_vals <- rgamma(num_spp, shape = 2, rate = 1 / (max_minus_min / 4)) + mean - max_minus_min / 2
  }
  return(rd_vals)
}

# Generate new dataset with different distributions
set.seed(123)
dd_new <- crossing(num_spp = c(2, 4, 8, 16),
                   max_minus_min = round(seq(1, 50, length = 45), 3),
                   mean = round(seq(-25, 25, length = 49), 3),
                   dist = c("normal", "uniform", "gamma")) %>%
  rowwise() %>%
  mutate(rd_vals = list(generate_rd_vals(num_spp, max_minus_min, mean, dist))) %>%
  unnest(cols = rd_vals)

# Calculate diversity metrics for new data
summary_stats_new <- dd_new %>%
  group_by(num_spp, max_minus_min, mean, dist) %>%
  summarise(
    dissimilarity = resp_div(rd_vals, sign_sens = FALSE),
    divergence = resp_div(rd_vals, sign_sens = TRUE),
    sign = sign(rd_vals),
    balance = abs(sum(rd_vals)),
    magnitude = sum(abs(rd_vals)),
    symmetry = mean(sign)
  )

```



```{r}
# Plot: Compare magnitude, balance, and symmetry across different distributions
summary_stats_new %>%
  group_by(num_spp, dist) %>%
  mutate(magnitude_rescaled = rescale(balance / num_spp)) %>%
  ggplot(aes(x = symmetry, y = magnitude, color = magnitude)) +
  geom_point() +
  facet_grid(num_spp ~ dist, scales = "free") +
  scale_color_viridis_c(option = "plasma") +
  labs(color = "Rescaled Balance", x = "Symmetry", y = "Magnitude") +
  theme_bw() +
  ggtitle("Effect of Different Distributions on Magnitude vs Symmetry")

```

Changing mean of the distribution

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)

# Updated function to generate rd_vals with different means and distributions
generate_rd_vals <- function(num_spp, max_minus_min, mean_shift, dist = "normal") {
  mean_val <- mean_shift  # Set the mean based on the input shift
  if (dist == "normal") {
    rd_vals <- rnorm(num_spp, mean = mean_val, sd = max_minus_min / 4)
  } else if (dist == "uniform") {
    rd_vals <- runif(num_spp, min = mean_val - max_minus_min / 2, max = mean_val + max_minus_min / 2)
  } else if (dist == "gamma") {
    rd_vals <- rgamma(num_spp, shape = 2, rate = 1 / (max_minus_min / 4)) + mean_val - max_minus_min / 2
  }
  return(rd_vals)
}

# Generate new dataset with different distributions and varying means
set.seed(123)
dd_updated <- crossing(num_spp = c(2, 4, 8, 16),
                       max_minus_min = round(seq(1, 50, length = 45), 3),
                       mean_shift = seq(-20, 20, length = 5),  # Varying mean from -10 to 10
                       dist = c("normal", "uniform", "gamma")) %>%
  rowwise() %>%
  mutate(rd_vals = list(generate_rd_vals(num_spp, max_minus_min, mean_shift, dist))) %>%
  unnest(cols = rd_vals)

# Calculate diversity metrics for the updated data
summary_stats_updated <- dd_updated %>%
  group_by(num_spp, max_minus_min, mean_shift, dist) %>%
  summarise(
    dissimilarity = resp_div(rd_vals, sign_sens = FALSE),
    divergence = resp_div(rd_vals, sign_sens = TRUE),
    sign = sign(rd_vals),
    balance = abs(sum(rd_vals)),
    magnitude = sum(abs(rd_vals)),
    symmetry = mean(sign)
  )


```



```{r}
# Plot: Compare magnitude, balance, and symmetry across different means and distributions
summary_stats_updated %>%
  group_by(num_spp, dist, mean_shift) %>%
  mutate(magnitude_rescaled = rescale(balance / num_spp)) %>%
  ggplot(aes(x = symmetry, y = magnitude, color = magnitude_rescaled)) +
  geom_point() +
  facet_grid(num_spp ~ dist, scales = "free", labeller = labeller(mean_shift = label_both)) +
  scale_color_viridis_c(option = "plasma") +
  labs(color = "Rescaled Balance", x = "Symmetry", y = "Magnitude") +
  theme_bw() +
  ggtitle("Effect of Different Distributions and Means on Magnitude vs Symmetry")

```

This expanded analysis helps explore whether the patterns of stability metrics hold under different statistical distributions and mean shifts, aligning with broader ecological interpretations of diversity and compensatory dynamics. The results provide insights into the robustness of these relationships across different scenarios, contributing to a more comprehensive understanding of the mechanisms underlying community stability.


```{r}
# Summarize the mean of rd_vals for each group
mean_check <- dd_updated %>%
  group_by(num_spp, dist, mean_shift) %>%
  summarise(mean_rd_vals = mean(rd_vals), .groups = "drop")

# Display the summary to check the means
print(mean_check)

```

```{r}
# Plot to show how the mean of rd_vals changes with mean_shift
ggplot(mean_check, aes(x = mean_shift, y = mean_rd_vals, color = dist)) +
  geom_point(size = 3) +
  geom_line() +
  facet_wrap(~num_spp, scales = "free") +
  labs(title = "Mean of rd_vals Across Different mean_shift Values",
       x = "Input mean_shift",
       y = "Calculated Mean of rd_vals",
       color = "Distribution") +
  theme_bw()

```

```{r}
summary_stats_updated %>% dplyr::filter(dist == "normal") %>% 
  group_by(num_spp, dist, mean_shift) %>%
  mutate(magnitude_rescaled = rescale(balance / num_spp)) %>%
  ggplot(aes(x = symmetry, y = magnitude, color = magnitude_rescaled)) +
  geom_jitter(width = 0.05) +
  facet_wrap(~ mean_shift) +
  scale_color_viridis_c(option = "plasma") +
  labs(color = "Rescaled Balance", x = "Symmetry", y = "Magnitude") +
  theme_bw() +
  ggtitle("Effect of Mean Shift on Magnitude vs Symmetry")

```

