---
title: "Analysis of the experiment looking at the effect of response diversity on community stability in fluctuating environments"
author: "Til Hämmig, Francesco Polazzo"
date: "2024-09-27"
output: html_document
---

```{r setup, include=FALSE}
library("ggplot2")
library("tidyverse")
library("gridExtra")
#library(rEDM)
library("parallel")
library("pracma")
library("purrr")
library("signal")
library("ggfortify")
library("data.table")
library("patchwork")
library("codyn")
library("ggrepel")
library("lme4")
library("lmerTest")
library("MuMIn")
library("RColorBrewer")
library("broom")
library("relaimpo")
library("lavaanPlot")
library(ggbeeswarm)
#devtools::install_github("canankarakoc/r_package_EDMhelper/EDMhelper")

```


```{r load packages, warning=FALSE, echo=FALSE}

divergence_df <- read_csv("Data/divergence_df.csv")
load("Data/dens_biomass_poly.RData")

dd_all_pred<-read.csv("Data/morph_dd_pred.csv")
dd_all_pred_nonoise<-read.csv("Data/morph_dd_pred_nonoise.csv")

load("Data/ciliate_traits.Rdata")

df_slopes <- read_csv("Data/df_slopes_cor.csv")
```

## Biomass

Let's have a look at the biomass dynamics in the different environmental treatments.

Is reaslised imbalance calculated in this chuck?

```{r biomass calcuations,echo=F, warning=FALSE}
# needs to have id_new variable
ciliate_traits <- ciliate_traits %>%
  mutate(
    # Remove dots from the date
    cleaned_date = gsub("\\.", "", date),
    # Extract the part of id after the underscore
    id_suffix = sub(".*_(.*)", "\\1", id),
    # Combine cleaned_date, id_suffix, and species_initial into a new variable
    id_new = paste0(cleaned_date, id_suffix, composition)
  ) %>%
  # Optionally, remove the intermediate columns to clean up
  dplyr::select(-cleaned_date, -id_suffix,-new_id)

uniqueN(ciliate_traits$id_new)==nrow(ciliate_traits) # all unique  ;)

id_dd<-full_join(dd_all_pred,dplyr::select(ciliate_traits,id_new,biomass),join_by("id_new"))


## add day variable

#create a day variable from the date variable

id_dd<-mutate(id_dd,date=as.Date(date,format = "%d.%m.%y"))

earliest_date<-min(id_dd$date)
days_since_earliest<-as.numeric(id_dd$date-earliest_date)+1
id_dd<-id_dd%>%mutate(day=days_since_earliest)

#create a summarised df on microcosm level with each species seperate
# Make sure, that we have n_frames and not N_frames
names(id_dd)[names(id_dd) == "N_frames"] <- "n_frames"

#extrapolation_factor <- 9.301902  # for 16 x magnification 
extrapolation_factor <- 9.828125  # for 25 x magnification 
video_biomass_species <- c( "C", "P", "S","D","L","T")

biomasses <- id_dd %>%
  group_by( day,temperature,nutrients,sample_ID,composition,predict_spec) %>% # group  by xxx
  summarize(
    biomass = sum(biomass * n_frames, na.rm = TRUE) / (1 * 125) # if not 3 videos corrections is done below with dens_factor
  ) %>%
  mutate(
    biomass = biomass * extrapolation_factor,
    )

biomasses<-biomasses%>%mutate(biomass=biomass*1000)

dd_ts_id<-biomasses

#fill up missing dates with biomass<-0

fill_dd<-expand.grid(sample_ID=unique(dd_ts_id$sample_ID),day=unique(dd_ts_id$day),predict_spec=unique(dd_ts_id$predict_spec))
complete_ts<-full_join(fill_dd,dd_ts_id,join_by(sample_ID,day,predict_spec))

complete_ts$biomass[is.na(complete_ts$biomass)]<-0
complete_ts<-complete_ts%>%mutate(composition=sub("_.*", "", sample_ID))
complete_ts<-complete_ts %>%
  mutate(temperature = sapply(strsplit(as.character(sample_ID), "_"), function(x) paste(x[3], x[4], sep = "-")))
complete_ts<- mutate(complete_ts,nutrients = gsub(".*Nut(.*?)_.*", "\\1", sample_ID))

# Now remove wrong combinations of composition and predict_spec / predict_spec

complete_ts<- complete_ts %>%
  rowwise() %>%
  dplyr::filter(predict_spec %in% unlist(strsplit(composition, ""))) %>%
  ungroup()  
complete_ts<-mutate(complete_ts,temperature=as.character(temperature),
                    nutrients=as.character(nutrients),
                    richness=nchar(composition))

complete_ts<-complete_ts%>%group_by(sample_ID,composition,day)%>%mutate(tot_biomass=sum(biomass))
complete_ts<-complete_ts%>%mutate(biom_contribution=biomass/tot_biomass)

complete_ts<-complete_ts%>%mutate(temperature=paste0(temperature," °C"),
                                      nutrients=paste0(nutrients," g/L"))


# introduce slopes of 
names(df_slopes)[names(df_slopes)=="species_initial"]<-"predict_spec"

slope_ts<-full_join(dplyr::select(df_slopes,nutrients,predict_spec,temperature,slope),complete_ts)
slope_ts<-slope_ts%>%mutate(w_slope=biom_contribution*slope,
                            sign=sign(slope))

slope_ts<-slope_ts%>%group_by(sample_ID,temperature,nutrients,richness,composition,day,tot_biomass)%>%summarize(
  sum_w_slopes=abs(sum(w_slope)),
                   mean_abs_slope=mean(abs(slope)),
  sum_abs_slope=sum(abs(slope)),
  abs_sum_slope=abs(sum(slope)),
  symmetry=abs(sum(sign)))


slope_ts<-slope_ts%>%mutate(richness=as.factor(richness))


##create new variable where it checks, where the last observation =0 is; with complete_ts
aggr_ts <- slope_ts %>%
  group_by( sample_ID) %>%
  arrange(day) %>%
  mutate(
    # Create a flag for non-zero tot_biomass
    non_zero_biomass = tot_biomass != 0,
    # Find the last non-zero day
    last_non_zero_day = ifelse(any(non_zero_biomass), max(day[non_zero_biomass], na.rm = TRUE), NA),
    # Find the first zero day after the last non-zero day
    first_zero_day = ifelse(
      !is.na(last_non_zero_day),
      min(day[!non_zero_biomass & day > last_non_zero_day], na.rm = TRUE),
      NA
    ),
    # Flag for days after the first zero day
    is_after_first_zero_day = ifelse(!is.na(first_zero_day), day > first_zero_day, FALSE)
  ) %>%
  ungroup()

aggr_ts<-aggr_ts%>%mutate(rep_var=sub("_[^_]+$", "", sample_ID))

biomass_ts<-aggr_ts%>%group_by(day,temperature,nutrients,richness)%>%summarize(tot_biom=mean(tot_biomass),se_tot_biom=sd(tot_biomass)/sqrt(as.numeric(length(tot_biomass))))


```

### tot biomass plot

```{r plot_biomass,echo=F, fig.cap='Community total biomass during the experiment in different environmental treatments. Different color represent richness levels.', fig.align="center", fig.height=8, fig.width=15}

plot_biomass<-ggplot(data=biomass_ts)+
  geom_ribbon(aes(x=day,y=tot_biom,fill=richness,color=richness,ymin=tot_biom-se_tot_biom,ymax=tot_biom+se_tot_biom),show.legend=F,alpha=0.5)+
  geom_line(method="loess",se=F,aes(x=day,y=tot_biom,color=richness),size=0.8)+
  scale_color_viridis_d(option = "inferno", begin = 0.2, end = 0.8)+
   scale_fill_viridis_d(option = "inferno", begin = 0.2, end = 0.8)+
  theme_classic()+
  #geom_text_repel( data=biomass_ts%>%dplyr::filter(day==58,tot_biom!=0),aes(x=day,y=tot_biom,group=composition,label = composition), size = 4,max.overlaps = 20) +
  facet_wrap(temperature~nutrients,scale="free_y")+
  theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        strip.background = element_rect(fill = "grey80", color = NA),  
    strip.text = element_text(size = 12))+
  ylab(expression("Community biomass [mg ml"^-1 * "]"))
  
plot_biomass
```

## Main Results 
```{r main results}
## discard all the rows that are part of time stretches where entire microcosms has (seemingly) gone extinct
#%>%dplyr::filter(is_after_first_zero_day==FALSE)

complete_aggr<-aggr_ts%>%group_by(composition,nutrients,temperature,sample_ID)%>%reframe(
  avg_w_sumslopes=mean(sum_w_slopes,na.rm = T),
  abs_sum_slope=mean(abs_sum_slope),
  magnitude=mean(mean_abs_slope),
  symmetry=mean(symmetry),
  sum_abs_slope=mean(sum_abs_slope),
  CV=sd(tot_biomass)/mean(tot_biomass),
  mean_biomass=mean(tot_biomass))

complete_aggr<-full_join(divergence_df,complete_aggr,by=join_by(composition,nutrients,temperature))%>%
  mutate(richness=as.factor(richness))




```

### Effect of T, N and R
```{r boxplots TNR, fig.cap=Effects of richness (a), temperature (b), and nutrients (c) on community total biomass temporal stability.', fig.align="center", fig.height=8, fig.width=15}}
#### effects of temperature; nutrient and richness, now with complete time series

plot_T_complete<-ggplot(data=complete_aggr)+
  geom_boxplot(aes(y=log10(1/CV),x=temperature))+
  geom_jitter(aes(y=log10(1/CV),x=temperature))

plot_T_complete<-ggplot(data=complete_aggr)+
  geom_quasirandom(data= complete_aggr,aes(y=log10(1/CV), x=temperature, group=temperature, colour=as.factor(temperature)),
                                  dodge.width=2, size=1) + 

  xlab("Temperature regime")+
  labs(color="Temperature regime", tag = "(b)")+
  theme(legend.position = "none",axis.title.y = element_blank())+
  scale_color_viridis_d(option = "inferno", begin = 0.2, end = 0.8)+
  scale_y_continuous(
    breaks = log10(c(0.1,0.2, 0.4, 0.8, 1.6, 3.2)),
    labels = c("0.1","0.2"," 0.4", "0.8", "1.6", "3.2")
  ) 


plot_N_complete<-ggplot(data=complete_aggr)+
  geom_boxplot(aes(y=log10(1/CV),x=nutrients))+
  geom_jitter(aes(y=log10(1/CV),x=nutrients,color=richness))

plot_N_complete<-ggplot(data=complete_aggr)+
  geom_quasirandom(data= complete_aggr,aes(y=log10(1/CV), x=nutrients, group=nutrients, colour=as.factor(nutrients)),
                                  dodge.width=2, size=1) + 

  xlab("Nutrients")+
  theme(legend.position = "none",axis.title.y = element_blank())+
  scale_color_viridis_d(option = "inferno", begin = 0.2, end = 0.8)+
  scale_y_continuous(
    breaks = log10(c(0.1,0.2, 0.4, 0.8, 1.6, 3.2)),
    labels = c("0.1","0.2"," 0.4", "0.8", "1.6", "3.2")
  ) + labs(tag = "(c)")

plot_R_complete<-ggplot(data=complete_aggr)+
  geom_boxplot(aes(y=log10(1/CV),x=richness))+
  geom_jitter(aes(y=log10(1/CV),x=richness))

plot_R_complete<-ggplot(data=complete_aggr)+
  geom_quasirandom(data= complete_aggr,aes(y=log10(1/CV), x=richness, group=richness, colour=as.factor(richness)),
                                  dodge.width=2, size=2) + 

  
  
  xlab("Richness")+
  theme(legend.position = "none",
        axis.title.y = element_blank())+
  scale_color_viridis_d(option = "inferno", begin = 0.2, end = 0.8)+
  scale_y_continuous(
    breaks = log10(c(0.1,0.2, 0.4, 0.8, 1.6, 3.2)),
    labels = c("0.1","0.2"," 0.4", "0.8", "1.6", "3.2")
  ) + labs (tag = "(a)")

combined_plot<-plot_R_complete/
(plot_T_complete+
plot_N_complete)

wrap_elements(combined_plot) +
  labs(tag = "Temporal stability\nlog(mean/s.d.)")+
  theme(
    plot.tag = element_text(size = rel(1), angle = 90),
    plot.tag.position = "left"
  )

# expression(atop("Temporal stability", log[10]("mean/s.d.")))



```


We can see that richness does not have a clear effect on community temporal stability, while stability was higher at lower temperature, and nutrients increased community temporal stability.

###Effect RD

We are now going to look at how response diversity (imbalance) affected temporal stability of total community biomass. We are going to look at the relationship between fundamental imbalance (so based only on species response surfaces measured in monoculture), an realised imbalance (measured accounting for species contribution to imbalance).

This is fundamentally testing our most important hypothesis.

```{r effect RD, fig.cap=Effects of fundamental and realised response diversity (measured as imbalance) on total community biomass temporal stability.', fig.align="center", fig.height=10, fig.width=16}

balance_dd<-complete_aggr%>%dplyr::select(abs_sum_slope,richness,CV)%>%
  cbind(type="fundamental")%>%mutate(abs_sum_slope=abs_sum_slope/as.numeric(richness))
names(balance_dd)[names(balance_dd)=="abs_sum_slope"]<-"balance"

weighted_balance_dd<-complete_aggr%>%dplyr::select(avg_w_sumslopes,richness,CV)%>%
  cbind(type="realised")
names(weighted_balance_dd)[names(weighted_balance_dd)=="avg_w_sumslopes"]<-"balance"

main_r_dd<-rbind(balance_dd,weighted_balance_dd)

plot_main <- main_r_dd %>%
  ggplot(aes(x = log10(balance), y = log10(1/CV), color = type)) +
  geom_point(aes(x = log10(balance), y = log10(1/CV))) +
  geom_smooth(method = "lm", aes(x = log10(balance), y = log10(1/CV)),show.legend=F) +
  theme_classic() +
  scale_color_viridis_d(option = "inferno", begin = 0.3, end = 0.6) +
  scale_x_continuous(
    breaks = log10(c(0.0025,0.01, 0.04, 0.16, 0.64)),
    labels = c("0.0025","0.01", "0.04", "0.16", "0.64")
  ) +
  scale_y_continuous(
    breaks = log10(c(0.1,0.2, 0.4, 0.8, 1.6, 3.2)),
    labels = c("0.1","0.2"," 0.4", "0.8", "1.6", "3.2")
  ) +
  ylab("Temporal stability\nlog(mean/s.d.)")+
  xlab("log(imbalance)") +
 theme(
    axis.title.x = element_text(size = 25),
    axis.title.y = element_text(size = 25),
    axis.text = element_text(size = 14),                      # Axis text size             # Facet label text size
    legend.title = element_text(size = 20),                   # Legend title size
    legend.text = element_text(size = 20),                     # Legend text size
        strip.background = element_rect(fill = "grey80", color = NA),  
    strip.text = element_text(size = 22)) +
  labs(color = "Imbalance") +  
  facet_wrap(~richness, labeller = as_labeller(function(x) paste0("richness = ", x)))

plot_main
```

We can see that imbalance is always negative related to temporal stability, which means that response diversity promotes stability across richness levels. Interestingly, we see that there is not difference in the sign of the relationship between fundamental and realised imbalance. Yet, as the richness increases, the relationship between realised imbalance and stability becomes steeper compared to fundamental imbalance. 


We look also at the relationship between divergence (our original response diversity metric)

```{r divergence_CV, fig.cap=Relationship between Divergence and temporal stability of total community biomass.', fig.align="center", fig.height=10, fig.width=16}

plot_CV_divergence<-complete_aggr%>%ggplot(aes(x=divergence,y=log10(1/CV),color=richness))+
  geom_point()+
  geom_smooth(method="lm",aes(x=divergence,y=log10(1/CV),color=richness),show.legend=F)+theme_classic()+
  scale_color_viridis_d(option = "inferno", begin = 0.2, end = 0.8)+
  scale_y_continuous(
    breaks = log10(c(0.1,0.2, 0.4, 0.8, 1.6, 3.2)),
    labels = c("0.1","0.2"," 0.4", "0.8", "1.6", "3.2")
  ) +
   theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        strip.background = element_rect(fill = "grey80", color = NA),  
    strip.text = element_text(size = 12)) +
  facet_wrap(~richness, labeller = as_labeller(function(x) paste0("richness = ", x)))+
  ylab("Temporal stability\nlog(mean/s.d.)")+
  xlab("divergence") 


plot_CV_divergence

```
We can see that the positive relationship between temporal stability and response diversity measured as divergence holds, but it becomes shallower as richness increases. We could speculated that this due to divergence considering only the responses of the 2 most extreme species. Thus, when species richness increases, disregarding the responses of the other species in the community except the 2 responding the most makes the relationship between response diversity and stability weaker. 


#Linear models

```{r}

lm_full<-lm(data=complete_aggr,log10(stability)~log10(abs_sum_slope)+richness+nutrients+temperature)
summary(lm_full)

lm_full_w<-lm(data=complete_aggr,log10(stability)~log10(avg_w_sumslopes)+richness+nutrients+temperature)
summary(lm_full_w)

lm_divergence<-lm(data=complete_aggr,log10(stability)~divergence*richness+nutrients+temperature)
summary(lm_divergence)

anova(lm_full,lm_divergence)

```




### sum vs. weighted_sum
```{r Cor sum weighted_sum}

richness2<-complete_aggr%>%dplyr::filter(richness==2)
cor_r2<-cor.test(log10(richness2$abs_sum_slope),log10(richness2$avg_w_sumslopes))$estimate

richness3<-complete_aggr%>%dplyr::filter(richness==3)
cor_r3<-cor.test(log10(richness3$abs_sum_slope),log10(richness3$avg_w_sumslopes))$estimate

richness4<-complete_aggr%>%dplyr::filter(richness==4)
cor_r4<-cor.test(log10(richness4$abs_sum_slope),log10(richness4$avg_w_sumslopes))$estimate


plot_cor<-ggplot(data=complete_aggr,aes(x=log10(abs_sum_slope),y=avg_w_sumslopes%>%log10(
)))+
  geom_point(aes(color=richness))+
   scale_y_continuous(
    breaks = log10(c(0.0025,0.01, 0.04, 0.16, 0.64)),
    labels = c("0.0025","0.01", "0.04", "0.16", "0.64")
  ) +
  scale_x_continuous(
    breaks = log10(c(0.01, 0.04, 0.16, 0.64)),
    labels = c("0.01", "0.04", "0.16", "0.64")
  ) +
   theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        strip.background = element_rect(fill = "grey80", color = NA),  
    strip.text = element_text(size = 12),legend.position="none") +
   geom_abline(slope = 1, intercept = 0, color = "black")+
   xlab(expression(log(balance)["fund"] )) +
  ylab(expression(log(balance)["real"] ))+
  facet_wrap(~richness, labeller = as_labeller(function(x) paste0("richness = ", x)))+
  scale_color_viridis_d(option = "inferno", begin = 0.2, end = 0.8)

plot_cor





```




# Asynchrony

```{r asynchrony calc.,echo=F}

#filter out completely extinct microcosms


async_df<-complete_ts%>%mutate(rep_var=sub("_[^_]+$", "", sample_ID))


async_Gross <- async_df %>% group_by(composition,sample_ID) %>%  
  do(synchrony_Gross = synchrony(., "day", "predict_spec", 
                              "biomass", metric = "Gross"))
async_Gross<-mutate(async_Gross,synchrony_Gross=synchrony_Gross%>%unlist())

async_Loroeau <- async_df %>% group_by(composition,sample_ID) %>%  
  do(synchrony_L = synchrony(., "day", "predict_spec", 
                              "biomass", metric = "Loreau"))
async_Loroeau<-mutate(async_Loroeau,synchrony_Loreau=synchrony_L%>%unlist())

async_aggr<-full_join(async_Gross,complete_aggr,join_by(sample_ID,composition))

async_aggr<-full_join(async_aggr,async_Loroeau,join_by(sample_ID,composition))

async_aggr_f<-async_aggr%>%dplyr::filter(nutrients!="0.01 g/L")
async_aggr<-async_aggr%>%mutate(rep_var=sub("_[^_]+$", "", sample_ID))
```


### Plot stability vs. Asynchrony Gross
```{r async plots,echo=F}
plot_asynch_CV_G<-ggplot(data=async_aggr,aes(x=-synchrony_Gross,y=log10(1/CV),color=nutrients))+
  geom_point()+
  geom_smooth(method="lm",show.legend=F)+
  theme_classic()+
  scale_y_continuous(
    breaks = log10(c(0.1,0.2, 0.4, 0.8, 1.6, 3.2)),
    labels = c("0.1","0.2"," 0.4", "0.8", "1.6", "3.2")
  ) +
   theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        strip.background = element_rect(fill = "grey80", color = NA),  
    strip.text = element_text(size = 12),legend.position="none") +
   ylab("Temporal stability\nlog(mean/s.d.)")+
  xlab("Asynchrony")+
  facet_wrap(~nutrients, labeller = as_labeller(function(x) paste0("nutrients = ", x)))+
  scale_color_viridis_d(option = "inferno", begin = 0.2, end = 0.8)




plot_asynch_CV_G

```


```{r async cor}

plot_asynch_CV_G<-ggplot(data=async_aggr,aes(x=log10(abs_sum_slope),y=-synchrony_Gross,color=nutrients))+
  geom_point()+
  geom_smooth(method="lm",show.legend=F)+
  theme_classic()+
  scale_x_continuous(
    breaks = log10(c(0.0025,0.01, 0.04, 0.16, 0.64)),
    labels = c("0.0025","0.01", "0.04", "0.16", "0.64")
  ) +
   theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        strip.background = element_rect(fill = "grey80", color = NA),  
    strip.text = element_text(size = 12),legend.position="none") +
   xlab(expression(log(balance)["fund"] )) +
  ylab("Asynchrony")+
  facet_wrap(~nutrients, labeller = as_labeller(function(x) paste0("nutrients = ", x)))+
  scale_color_viridis_d(option = "inferno", begin = 0.2, end = 0.8)

plot_asynch_CV_G

```









# SEM and balance sim

```{r interaction RD async}

# simulate random slopes


slopes<-rnorm(1000,mean=0,sd=0.1)

mean_s<-mean(df_slopes$slope)
sd_s<-sd(df_slopes$slope)

RD_df<-lapply(c(2,3,4,10),function(r){
  
  RD<-lapply(1:1000,function(c){
  slopes<-rnorm(r,mean=0,sd=sd_s)
  balance<-abs(mean(slopes))
  divergence<-(max(slopes)-min(slopes)-abs(abs(max(slopes))-abs(min(slopes))))/(max(slopes)-min(slopes))
  adj_divergence<-abs(sum(slopes))/sum(abs(slopes))
  avg_magnitude<-mean(abs(slopes))
  RD_df<-data.frame(balance=balance,divergence=divergence,richness=r,adj_divergence=1-adj_divergence,avg_magnitude=avg_magnitude)
  
  return(RD_df)
})
  RD<-do.call("rbind",RD)
})

RD_df<-do.call("rbind",RD_df)


plot<-ggplot(data=RD_df,aes(x=balance,color=as.factor(richness)))+
  geom_density()

plot_sim_div_bal<-ggplot(data=RD_df,aes(x=adj_divergence,y=balance,color=avg_magnitude))+
  geom_point()+
  facet_wrap(~richness)






# Ensure the data is ungrouped before applying transformations
sem_aggr <- async_aggr %>%
  ungroup() %>%  # Ensure there is no grouping
  mutate(
    log_abs_sum_slope = log10(abs_sum_slope),
    stability = log10(1 / CV),
    richness = as.numeric(richness),
    temperature=temperature,
    log_magnitude=log10(magnitude),
    sqrt_symm=sqrt(symmetry),
    log_sum_abs=log10(sum_abs_slope),
    RD=abs_sum_slope/sum_abs_slope,
    log_syn_Gross=log10(1+synchrony_Gross),
    symmetry_scaled=(symmetry/as.numeric(richness)),
    asyn_Gross= (-synchrony_Gross),
    log_asyn_Gross=log10((-synchrony_Gross)+1),
    log_sum_abs_slope=log10(sum_abs_slope)
    #Keep it as an ordered factor
  )

quantile(sem_aggr$log_abs_sum_slope)


sem_aggr$nutrients[sem_aggr$nutrients=="0.01 g/L"]<-0.01
sem_aggr$nutrients[sem_aggr$nutrients=="0.35 g/L"]<-0.35
sem_aggr$nutrients[sem_aggr$nutrients=="0.75 g/L"]<-0.75

sem_aggr$temperature[sem_aggr$temperature=="18-21 °C"]=1
sem_aggr$temperature[sem_aggr$temperature=="22-25 °C"]=2
sem_aggr$temperature[sem_aggr$temperature=="25-28 °C"]=3


#correct balance for magnitude
# the effect of magnitude goes down, the more divergence there is

# Load the lavaan package
library(lavaan)


model1A <- '
  stability ~ asyn_Gross+temperature+nutrients+symmetry_scaled+log_sum_abs_slope
  
  asyn_Gross ~ nutrients+richness+log_sum_abs_slope+symmetry_scaled


'
modl<-lm(data=sem_aggr,log_abs_sum_slope~symmetry_scaled+log_sum_abs_slope)


model1B <- '
  stability ~ asyn_Gross
  +temperature
  +nutrients
  +log_abs_sum_slope+
  richness

  
  asyn_Gross ~ log_abs_sum_slope
  +nutrients
  +richness
'



model1C <- '
  stability ~ asyn_Gross+temperature+c("c1","c2","c1")*richness+log_abs_sum_slope
  
  asyn_Gross ~ log_abs_sum_slope+c("d1","d2","d3")*richness
  
  c1==0
  d2==0
'




# Define the SEM model using lavaan syntax
 model2 <- '
  # Direct effects
  stability ~ asyn_Gross+log_magnitude
  
  # Indirect effect via synchrony
  asyn_Gross ~ symmetry


 
'
 


# Fit the model
fit <- sem(model2, data = sem_aggr%>%dplyr::filter(!is.na(synchrony_Gross),nutrients!="0.01 g/L"))


fit1A <- sem(model1A, estimator="MLM",meanstructure = TRUE,data = sem_aggr%>%dplyr::filter(!is.na(synchrony_Gross)))


fit1B <- sem(model1B, estimator="MLM",meanstructure = TRUE,data = sem_aggr%>%dplyr::filter(!is.na(synchrony_Gross)))

fit1C <- sem(model1C, estimator="MLM",meanstructure = TRUE,group="nutrients", data = sem_aggr%>%dplyr::filter(,!is.na(synchrony_Gross)))



 ### model B slightly better
AICc(fit1C,fit1B) ### model B and C almost the same

anova(fit1Ba,fit1Bb)
# Summarize the results
summary(fit1B, standardized = TRUE,rsquare=T, fit.measures = TRUE)


modificationindices(fit1B)


lavaanPlot(model = fit1A, node_options = list(shape = "box", fontname = "Helvetica"), edge_options = list(color = "grey"), coefs = TRUE, sig = .05)

```

# interactions

```{r stat averaging}

interactions_df <- read_csv("interactions_df_w_T.csv")

int_aggr<-full_join(interactions_df,async_aggr,join_by(sample_ID))
int_aggr$mean_int_strength[is.na(int_aggr$mean_int_strength)]=0
int_aggr$mean_interaction[is.na(int_aggr$mean_interaction)]=0
int_aggr$sum_int_strength[is.na(int_aggr$sum_int_strength)]=0

M_int<-lm(data=int_aggr,log10(stability)~log10(abs_sum_slope)+nutrients+temperature+richness+mean_int_strength)
anova(M_int,lm_full)



plot_int_syn<-ggplot(data=int_aggr,aes(y=synchrony_Gross,x=mean_int_strength,color=log10(abs_sum_slope)))+
  geom_point()+
  geom_smooth(method="lm")





### above rep level
rep_aggr<-int_aggr%>%group_by(rep_var.y,richness,temperature,nutrients,composition)%>% dplyr::reframe(log_mean_fun_balance=log10(mean(abs_sum_slope)),
                                                                                     mean_asyn_Gross=mean(-synchrony_Gross,na.rm=T),
                                                                                     mean_int_rep_strength=mean(mean_int_strength),
                                                                                     log_mean_stab=log10(mean(1/CV)))
                                                                              

M_int<-lm(data=rep_aggr,log_mean_stab~log_mean_fun_balance+nutrients+temperature+richness+mean_int_rep_strength)
anova(M_int,lm_full)


M_int_syn<-lm(data=rep_aggr,mean_asyn_Gross~log_mean_fun_balance+richness+temperature+mean_int_rep_strength+nutrients)
summary(M_int_syn)

plot_richness_int<-ggplot(data=rep_aggr,aes(x=mean_int_rep_strength))+
  geom_histogram(bins=10)+
  facet_wrap(~nutrients)
plot_int_syn<-ggplot(data=rep_aggr,aes(y=mean_asyn_Gross,x=mean_int_rep_strength))+
  geom_point()+
  geom_smooth(method="lm")+
  facet_wrap(~richness)

### sem

rep_aggr<-rep_aggr%>%mutate(richness=as.numeric(richness))




model_int <- '
  log_mean_stab ~ mean_asyn_Gross
  +temperature
  +nutrients
  +log_mean_fun_balance
  +richness


  
  
  mean_asyn_Gross ~ log_mean_fun_balance
  +nutrients
  +mean_int_rep_strength+
  +richness
  

  
'
fit_int <- sem(model_int, estimator="MLM",meanstructure = TRUE,data = rep_aggr)


summary(fit_int, standardized = TRUE, fit.measures = TRUE,rsquare=T)

AICc(fit_int)

lm_r_int_syn<-lm(data=rep_aggr,mean_asyn_Gross~richness+mean_int_rep_strength+log_mean_fun_balance)


##### load interactions from rgular CCM


interactions_df <- read_csv("interactions_df_regular.csv")

int_aggr2<-full_join(interactions_df,async_aggr,join_by(sample_ID))
int_aggr2$mean_interaction_strength[is.na(int_aggr2$mean_interaction_strength)]=0
int_aggr2$mean_sum_interactions[is.na(int_aggr2$mean_sum_interactions)]=0


M_int<-lm(data=int_aggr2,log10(stability)~log10(abs_sum_slope)+nutrients.y+temperature.y+richness+mean_interaction_strength)
anova(M_int,lm_full)

```

## Plot interaction vs Asnychrony


```{r}

```





##Poster

```{r}

### time series with nutrients==0.35 and T=22-25

poster_ts<-complete_ts%>%filter(nutrients=="0.35 g/L",
                                temperature=="22-25 °C",
                                sample_ID=="DLPS_Tmp_22_25_Nut0.35_3")
dark2_palette <- brewer.pal(n = 8, name = "Dark2")
selected_colors <- dark2_palette[2:5]

plot_poster_ts <- ggplot(data = poster_ts, aes(x = day, y = biomass, color = predict_spec)) +
  geom_line(linewidth=1.5) + ylab("Biomass")+xlab("Day")+
  theme_classic() +
  scale_color_manual(values = selected_colors) +
  theme(legend.position = "none",       
        axis.text.y = element_blank(),        # Remove the y-axis 
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size=16),
        axis.title.y=element_text(size=16))     

plot_poster_ts

cake_df<-poster_ts%>%group_by(predict_spec)%>%summarize(mean_biom=mean(biomass))


cake_df <- tibble::tibble(
  predict_spec = c("Dexiostoma", "Loxocephalus", "Paramecium", "Spirostomum"),
  mean_biom = c(0.000000843, 0.000000808, 0.00000520, 0.000000237)
)

# percentages
cake_df <- cake_df %>%
  mutate(percentage = mean_biom / sum(mean_biom) * 100)

# cake diagramm
plot_cake <- ggplot(cake_df, aes(x = "", y = percentage, fill = predict_spec)) +
  scale_fill_manual(values = selected_colors) +
  geom_bar(stat = "identity", width = 1) + 
  coord_polar(theta = "y") +                    # Titel und Legende
  theme_void()+
  labs(fill="Species")+
  theme(legend.direction = "vertical",legend.position="right",
        legend.title = element_text(size=16))# Verwende eine Farbpalette


plot_cake+plot_poster_ts+ plot_layout(ncol = 2,widths=c(1,2))   # Title for the combined plot
weight_plot <- wrap_elements(plot_cake + theme(plot.margin = margin( b = 150))) +  # Move the cake plot up
  plot_spacer() +
  plot_poster_ts + 
  plot_layout(ncol = 3, widths = c(1, 0.2, 1))

weight_plot

#ggsave("weight.pdf", plot = weight_plot, width = 10, height = 6)


# create temperature regime plot

regime1<-data.frame(temperature=rep(c(18,18,18,21,21,21),length.out=60),regime= "18-21 °C")
regime2<-data.frame(temperature=rep(c(22,22,22,25,25,25),length.out=60),regime= "22-25 °C")
regime3<-data.frame(temperature=rep(c(25,25,25,28,28,28),length.out=60),regime= "25-28 °C")
day=

df_Temp<-rbind(regime1,regime2,regime3)
df_Temp<-cbind(df_Temp,day=rep(1:60))

plot_temp<-ggplot(data=df_Temp,aes(x=day,y=temperature,linetype=regime))+
  geom_line()+  theme_classic() +
  ylab("Temperature °C")+xlab("Day")+
  labs(linetype="Temperature regime")+
  scale_y_continuous(breaks = c(28, 25, 22, 21, 18)) +
  theme(legend.direction = "vertical",legend.position="right",        
        axis.title.x = element_text(size=16),
        axis.title.y=element_text(size=16))


ggsave("temp_regime.pdf", plot = plot_temp, width = 8, height = 6)


############ Main results plot


balance_dd<-complete_aggr%>%dplyr::select(abs_sum_slope,richness,CV)%>%
  cbind(type="no")%>%mutate(abs_sum_slope=abs_sum_slope/as.numeric(richness))
names(balance_dd)[names(balance_dd)=="abs_sum_slope"]<-"balance"

weighted_balance_dd<-complete_aggr%>%dplyr::select(avg_w_sumslopes,richness,CV)%>%
  cbind(type="yes")
names(weighted_balance_dd)[names(weighted_balance_dd)=="avg_w_sumslopes"]<-"balance"

main_r_dd<-rbind(balance_dd,weighted_balance_dd)

plot_main <- main_r_dd %>%
  ggplot(aes(x = log10(balance), y = log10(1/CV), color = type)) +
  geom_point(aes(x = log10(balance), y = log10(1/CV))) +
  geom_smooth(method = "lm", aes(x = log10(balance), y = log10(1/CV))) +
  theme_classic() +
  scale_color_viridis_d(option = "inferno", begin = 0.3, end = 0.6) +
  scale_x_continuous(
    breaks = log10(c(0.0025,0.01, 0.04, 0.16, 0.64)),
    labels = c("0.0025","0.01", "0.04", "0.16", "0.64")
  ) +
  ylab(expression("temporal stability: "~log[10]("mean/s.d.")))+
  xlab(expression(log[10]("Balance"))) +
  theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        strip.background = element_rect(fill = "grey80", color = NA),  
    strip.text = element_text(size = 12)) +
  labs(color = "Weighted") +  
  facet_wrap(~richness, labeller = as_labeller(function(x) paste0("richness = ", x)))

ggsave("main_plot.pdf", plot = plot_main, width = 10, height = 6)



### linear model summary table

# Install and load the broom package if you haven't already

library(broom)

# Fit the model
lm_full<-lm(data=complete_aggr, log10(stability)~log10(abs_sum_slope)+richness+temperature+nutrients)

# Tidy up the summary
tidy_model <- tidy(lm_full_w)

# View the tidied model
tidy_model


library(broom)
library(writexl)



# Step 3: Tidy the model output using broom
model_summary <- tidy(lm_full)

# Step 4: Add significance stars for p-values
model_summary <- model_summary %>%
  mutate(Significance = case_when(
    p.value < 0.001 ~ "***",
    p.value < 0.01 ~ "**",
    p.value < 0.05 ~ "*",
    p.value < 0.1 ~ ".",
    TRUE ~ ""
  ))

# Step 5: Rename columns for a nicer presentation
names(model_summary) <- c("Term", "Estimate", "Std. Error", "t value", "p value", "Significance")

# Step 6: Write the summary to an Excel file
write_xlsx(model_summary, "full_model_summary.xlsx")


# Step 3: Tidy the model output using broom
model_summary <- tidy(lm_full_w)

# Step 4: Add significance stars for p-values
model_summary <- model_summary %>%
  mutate(Significance = case_when(
    p.value < 0.001 ~ "***",
    p.value < 0.01 ~ "**",
    p.value < 0.05 ~ "*",
    p.value < 0.1 ~ ".",
    TRUE ~ ""
  ))

# Step 5: Rename columns for a nicer presentation
names(model_summary) <- c("Term", "Estimate", "Std. Error", "t value", "p value", "Significance")

# Step 6: Write the summary to an Excel file
write_xlsx(model_summary, "full_model_w_summary.xlsx")



```

